Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based study

### 2024-02-27 manuscript.Rmd
27/2

This was a registry-based study using all patients between 2014 and 2022
from the Karolinska University Hospital in Solna who had been admitted
to the ICU and who were reviewed regarding the presence of OFI by the
mortality and morbidity conference. Bi-variable and multivariable
logistic regression was used to assess the following patient and process
factors and OFI among this group of patients: gender, age, survival
after 30 days, Glasgow Coma Scale (GCS), respiratory rate, systolic
blood pressure, working hours, weekend, time from arrival at the
hospital until first computed tomography (CT), if the patient was
intubated and Injury Severity Score (ISS).

@@

We conducted a retrospective single-centre cohort study based on data
from the Karolinska University Hospital in Solna and from the Swedish
Trauma registry. We used R for all statistical analyses [@19]. A
complete case analysis was conducted after handling missing values in
systolic blood pressure, respiratory rate and GCS score. We present
sample characteristics using descriptive statistics.

@@

The trauma registry includes all patients admitted with trauma team
activation, regardless of Injury Severity Score (ISS), as well as
patients admitted without trauma team activation but found to have an
ISS of more than 9. We included all patients who had been included in
the morbidity and mortality screening process between January 1, 2014
and June 1, 2022 and who were also admitted to the ICU. We excluded
patients who were younger than 15 years and patients who were dead on
arrival.

@@

The quantitative variables were factors chosen from the trauma registry,
based on the locally used audit filters, standard epidemiological
factors and factors registered in the Swedish Trauma Registry. The
categorical factors were sex, survival after 30 days, GCS, respiratory
rate, systolic blood pressure, working hours, weekend, time from arrival
at the hospital until first CT, if the patient was intubated, age and
Injury Severity Score (ISS). Working hours was defined as between 8.00
a.m. and 5 p.m., or during a weekend, defined as Saturday or Sunday.

@@

There were ... patients in the trauma registry. ... with trauma team
activation regardless of ISS score, ... patients admitted without trauma
team activation but with ISS score \> 9. .... patients were included in
the morbidity and mortality screening process between 2014 and 2022. ...
of these patients were also admitted to the ICU. ... patients were
excluded as they were younger than 15 years and ... patients were
excluded because they were dead on arrival. This leaves us with 167
outcomes.

@@

[n] patients were included in the trauma registry and trauma care
quality database between 2017 and 2022. [n] patients were excluded,
leaving a total of [n] patients eligible for the study. [n] were
excluded because they were under 15 years old. [n] patients were
excluded because they were dead on arrival. [n] were excluded because
they were not treated in the ICU, and [n] patients were excluded because
they had missing data (?).

[n{%}] unique patients of the [n] patients included were flagged with
OFI. Table 1 illustrates the number of patients flagged per audit filter
used by the Karolinska University Hospital in Solna.

Table 2 present sample characteristics. [n{%}] patients were male and
[n{%}] patients were female. [n{%}] patients had no intubation. Most
patients of the intubated ones were intubated for xx days [n{%}]. The
number of patients who arrived outside working hours - on Saturday or
Sunday or after 5 pm or before 8 am were [n{%}]. [n{%}] patients had an
ASA-score of 1-2 before arrival to the hospital, and [n{%}] patients had
an ASA-score of 3-7 before arrival to the hospital. [n{%}] patients died
within 30 days after hospitalization.

Table 3 shows the unadjusted and adjusted associations of the selected
patient and process factors in relation to OFI. In the unadjusted
analysis the factors that were significantly associated with the
presence of OFI were: ... Male
`r inline_text(table3a, variable = Sex, level = "Male")` Female
`r inline_text(table3a, variable = Sex, level = "Female")` Age
`r inline_text(table3a, variable = Age)` No intubation
`r inline_text(table3a, variable = Intubation, level = "No intubation")`
Respiratory rate `r inline_text(table3a, variable = RespiratoryRate)`
Systolic blood pressure
`r inline_text(table3a, variable = SystolicBloodPressure)` Glasgow Coma
Scale `r inline_text(table3a, variable = GlasgowComaScale)` Jour
`r inline_text(table3a, variable = Jour, level = "Yes")` Time to first
CT `r inline_text(table3a, variable = TimeFCT)` Time to first
intervention `r inline_text(table3a, variable = TimeFInt)` ASA preinjury
1-2 `r inline_text(table3a, variable = ASApreinjury, level = "ASA 1-2")`
Death within 30 days
`r inline_text(table3a, variable = Survival, level = "Dead")`

@@

```{=html}
<!-- (a) Report numbers of individuals at each stage of study—eg numbers
potentially eligible, examined for eligibility, confirmed eligible,
included in the study, completing follow-up, and analysed

(b) Give reasons for non-participation at each stage

(c) Consider use of a flow diagram

-->
```
## Descriptive data

```{=html}
<!-- (a) Give characteristics of study participants (eg demographic,
clinical, social) and information on exposures and potential
confounders

(b) Indicate number of participants with missing data for each
variable of interest

(c) Cohort study — Summarise follow-up time (eg, average and total
amount)
-->
```
## Outcome data

```{=html}
<!-- Cohort study — Report numbers of outcome events or summary measures
over time
-->
```
## Main results

```{=html}
<!-- (a) Give unadjusted estimates and, if applicable, confounder-adjusted
estimates and their precision (eg, 95% confidence interval). Make
clear which confounders were adjusted for and why they were included

(b) Report category boundaries when continuous variables were
categorized

(c) If relevant, consider translating estimates of relative risk into
absolute risk for a meaningful time period
-->
```
## Other analyses

```{=html}
<!--Report other analyses done—eg analyses of subgroups and interactions,
and sensitivity analyses 
-->
```
# Discussion

```{=html}
<!--

Key results
-----------
Summarise key results with reference to study objectives

Limitations
-----------

Discuss limitations of the study, taking into account sources of
potential bias or imprecision.  Discuss both direction and magnitude
of any potential bias

Interpretation
--------------

Give a cautious overall interpretation of results considering
objectives, limitations, multiplicity of analyses, results from
similar studies, and other relevant evidence

Generalisability
----------------

Discuss the generalisability (external validity) of the study results

-->
```
# Conclusion

>>>>

This was a registry-based study using all patients between 2014 and 2022
from the Karolinska University Hospital in Solna who had been admitted
to the ICU and who were reviewed regarding the presence of OFI by the
mortality and morbidity conference. Bi-variable and multivariable
logistic regression was used to assess the following patient and process
factors and OFI among this group of patients: gender, age, if the
patient was intubated or not and how many days, respiratory rate,
systolic blood pressure, Glasgow Coma Scale (GCS), Injury Severity Score
(ISS), on call, time from arrival at the hospital until first computed
tomography (CT), ASA preinjury and survival after 30 days.

@@

We conducted a retrospective single-centre cohort study based on data
from the Karolinska University Hospital in Solna and from the Swedish
Trauma registry. We used R for all statistical analyses [@19]. For
patients where hospital values for systolic blood pressure, respiratory
rate or GCS score were missing, prehospital values were used instead. A
complete case analysis was conducted after handling missing values. We
present sample characteristics using descriptive statistics.

@@

The trauma registry includes all patients admitted with trauma team
activation, regardless of Injury Severity Score (ISS), as well as
patients admitted without trauma team activation but found to have an
ISS of more than 9. We included all patients who had been included in
the morbidity and mortality screening process between January 1, 2014
and February 1, 2023 and who were also admitted to the ICU. We excluded
patients who were younger than 15 years and patients who were dead on
arrival.

@@

The quantitative variables were factors chosen from the trauma registry,
based on the locally used audit filters, standard epidemiological
factors and factors registered in the Swedish Trauma Registry. The
categorical factors were sex, age, if the patient was intubated or not
and how many days, respiratory rate, systolic blood pressure, Glasgow
Coma Scale (GCS), Injury Severity Score (ISS), on call, time from
arrival at the hospital until first computed tomography (CT), ASA
preinjury and survival after 30 days. On duty was defined as before 8.00
a.m., after 5 p.m., or during a weekend - defined as Saturday or Sunday.

@@

There were ... patients in the trauma registry between January 1 2014
and February 1 2023. ... with trauma team activation regardless of ISS
score, ... patients admitted without trauma team activation but with ISS
score \> 9. .... patients were included in the morbidity and mortality
screening process between 2014 and 2022. ... of these patients were also
admitted to the ICU. ... patients were excluded as they were younger
than 15 years and ... patients were excluded because they were dead on
arrival. This leaves us with 167 outcomes.

@@

14,022 patients were included in the trauma registry and trauma care
quality database between 2017 and 2023. 12,278 patients were excluded,
leaving a total of 1,744 patients eligible for the study. 11,343 were
excluded because they were not treated in the ICU. 3 patients were
excluded because they were under 15 years old. 6 patients were excluded
because they were dead on arrival, and 926 patients were excluded
because there was no data on the presence or absence OFI.

[n] patients were excluded because there was missing data, leaving a
total of [n] patients included in the study.

Table 1 present sample characteristics. [n{%}] patients were male and
[n{%}] patients were female. [n{%}] patients had no intubation. Most
patients of the intubated ones were intubated for xx days [n{%}]. The
number of patients who arrived outside working hours - on Saturday or
Sunday or after 5 pm or before 8 am were [n{%}]. [n{%}] patients had an
ASA-score of 1-2 before arrival to the hospital, and [n{%}] patients had
an ASA-score of 3-7 before arrival to the hospital. [n{%}] patients died
within 30 days after hospitalization.

Table 2 shows the unadjusted and adjusted associations of the selected
patient and process factors in relation to OFI. In the unadjusted
analysis the factors that were significantly associated with the
presence of OFI were: ... Male
`r inline_text(table3a, variable = Sex, level = "Male")` Female
`r inline_text(table3a, variable = Sex, level = "Female")` Age
`r inline_text(table3a, variable = Age)` No intubation
`r inline_text(table3a, variable = Intubation, level = "No intubation")`
Respiratory rate `r inline_text(table3a, variable = RespiratoryRate)`
Systolic blood pressure
`r inline_text(table3a, variable = SystolicBloodPressure)` Glasgow Coma
Scale `r inline_text(table3a, variable = GlasgowComaScale)` Jour
`r inline_text(table3a, variable = Jour, level = "Yes")` Time to first
CT `r inline_text(table3a, variable = TimeFCT)` Time to first
intervention `r inline_text(table3a, variable = TimeFInt)` ASA preinjury
1-2 `r inline_text(table3a, variable = ASApreinjury, level = "ASA 1-2")`
Death within 30 days
`r inline_text(table3a, variable = Survival, level = "Dead")`

@@

<!-- Flowchart -->

## Descriptive data

<!-- Sample characteristics - Table1, do I need to display number of missing for each variable? -->

## Main results

<!--  Adjusted and unadjusted analysis - Table 2 -->

# Discussion

### Key results

We found that --- was associated with higher odds of OFI. Previous
research ...

These findings could be because ...

\-- was associated with OFI in the unadjusted analysis but not in the
adjusted analysis.

1.  add sources from previous studies

2.  studies to back up the results from the analyses

### Limitations

There were several limitations in our study. The selection for review
for the morbidity and mortality review relied mostly on the local audit
filters used at the Karolinska Institute Hospital in Solna. This leads
to a risk of misplacement where some patients with OFI were not selected
by for the review based on the audit filters although there was an OFI.

Another limtation is that some relevant factors, such as ----, were not
part of the Swedish Trauma registry and therefore not adjusted for in
the analysis.

The study was a single-center study and illustrate the situation in
Stockholm. More study is needed on the area of OFI in patients admitted
to the ICU. A suggestion would be to conduct a multi-center study on
this group of patients to further investigate areas for improvement and
to further develop trauma quality programs.

Future studies could also focus on ---- which was seen to be associated
with OFI in these patients.

<!--Lägg till 2 st förslag på explicita fortsättningsstudier! -->

### Generalisability

The study is conducted on --- outcomes. There were few outcomes because
we included only those who were admitted to the ICU, \> 14 years old,
not dead on arrival and had a registration for the presence or absence
of OFI. The external validity is therefore compromised due to the
limited size of the Swedish database. However, ...

# Conclusion

---- among patients admitted to the ICU have higher odds of OFI. These
factors can be seen as new targets for further reseach to further
supplement trauma quality improvement programs.
